---
title: "Set 5"
author: "Oluwanifemi"
date: "2024-10-24"
output: pdf_document
---

```{r}
library(ggplot2)
library(stats)
library(readxl)

```


## Question One

```{r}


birth_dat <- read_excel("birth_dat.xlsx")
head(birth_dat) 
X <- model.matrix(~ PerCapIncome + PopFarms + InfantMort, data= birth_dat)
head(X)

y = birth_dat$Birthrate
XtX <- t(X) %*% X
beta.hat <- solve(XtX, t(X)%*%y)
beta.hat

model <- lm(Birthrate ~ PerCapIncome + PopFarms + InfantMort, data = birth_dat)  
coef(model) 
#The results are the similar so in comparison, they are equal to each other.

mexico_row <- birth_dat[birth_dat$Nation == "Mexico", ]
mexico_row
X_mexico <- model.matrix(~ PerCapIncome + PopFarms + InfantMort, data = mexico_row)
fitted_mexico <- X_mexico %*% beta.hat 
fitted_mexico 



new_country <- data.frame(
  PerCapIncome = 600,
  PopFarms = 0.2,
  InfantMort = 20
)

predicted_value <- predict(model, newdata = new_country)
predicted_value

residuals <- residuals(model)
RSS <- (sum(residuals^2))
RSS


```

## Question 2


```{r}
x <- 1:20
set.seed(463)
y <- x + rnorm(20)


model1 <-lm(y~x)
beta_model1 <- coef(model1)
beta_model1

X <- cbind(1, x)
XtX <- t(X) %*% X
beta_hat <- solve(XtX, t(X)%*%y)
beta_hat

x_2<- x^2
model2 <-lm(y~x + x_2)
coef(model2)
X <- model.matrix(~ x + x_2)
#head(X)
XtX1 <- t(X) %*% X
beta_hat1 <- solve(XtX1, t(X)%*%y)
beta_hat1



# Function to calculate estimates for polynomial terms up to degree 10
calculate_polynomial_estimates <- function(degree) {
 
  X_poly <- sapply(0:degree, function(d) x^d)
  
  # Attempting direct inversion
  tryCatch({
    XtX1 <- t(X_poly) %*% X_poly
    beta_direct <- solve(XtX1, t(X_poly)%*%y)
    return(beta_direct)
  }, error = function(e) {
    return("Direct inversion failed")
  })
}


for (degree in 0:10) {
  cat("Degree:", degree, "\n")
  print(calculate_polynomial_estimates(degree))
}

#It fails when the degree is 6.


set.seed(463)
x2 <- x + rnorm(20, sd = 0.1)
model3 <- lm(y ~ x + x2)
correlation_x_x2 <- cor(x, x2)
beta_lm3 <- coef(model3)
correlation_x_x2 
beta_lm3  
#Both of them are highly correlated


set.seed(463)
x3 <- x + rnorm(20, sd = 0.001)
model4 <- lm(y ~ x + x3)
beta_lm4 <- coef(model4)
cor(x, x3)  

beta_lm4

```
### c) It fails when the number of degree is 6.

### d) They are highly correlated (although this might raise the concern for multicollinearity).

### e) Using the correlation being "1", there might be an issue of multicollinearity.The coefficent for x is approximately -999 and the coefficent for x3 is approximately 1000, htis might show that an increase in x csn be an almost exactly offset by an increase in x3, btu it mgith seems the model is assigning high sensitivity for small chnages in data.
```{r}

```

