---
title: "Set_4"
author: "Oluwanifemi"
date: "2024-10-18"
output: pdf_document
---

```{r}
library(ggplot2)
library(stats)
library(readxl)

```



```{r}
 # Read the .xlsx file
data <- read_excel("birth_dat.xlsx")

# Check the first few rows of the data
head(data)
```

```{r}

model <- lm(Birthrate ~ PerCapIncome+PopFarms+InfantMort, data = data)

summary(model)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


### n is the number of observation
### p is number of predictor variables including intercept

```{r}
n <- nrow(data)
p <- 3
cat("n is", n, "and p is", p)


```
The underlying mean function for predicting the birth rate can be expressed as:

\[
\text{BirthRate} = \beta_0 + \beta_1 \times \text{PerCapIncome} + \beta_2 \times \text{PopFarms} + \beta_3 \times \text{InfantMort} + \epsilon
\]



```{r}

team <- head(data, 3)  # Get the first 3 rows

# Convert the Birthrate column into a 3x1 data frame and print
birthrate_df <- data.frame(Birthrate = team$Birthrate)
print(birthrate_df)

# Print the rest of the columns (excluding 'Birthrate')
team_without_birthrate <- team[, !(names(team) %in% "Birthrate")]
print(team_without_birthrate)



```


```{r}

x_9 <- data[9, ]
x_9_without_birthrate <- x_9[, !(names(x_9) %in% "Birthrate")]
print(x_9_without_birthrate)

x2 <- data$PerCapIncome[1:3]
print(x2)

subset_data <- data.frame(
  Nation = c("Venezuela", "Mexico", "Ecuador"),
  PerCapIncome = c(392, 118, 44)
)


print(subset_data)


```


## Sum of Squared Errors (SSE)

The equation we are minimizing can be expressed as:

\[
\text{SSE} = \sum_{i=1}^{n} (y_i - \hat{y}_i)^2
\]


## Linear Model

Given the linear model:

\[
\hat{y}_i = \beta_0 + \beta_1 \times \text{PerCapIncome}_i + \beta_2 \times \text{PopFarms}_i + \beta_3 \times \text{InfantMort}_i
\]



## Expanded Sum of Squared Errors

The sum of squared errors can be expanded as follows:

\[
\text{SSE} = \sum_{i=1}^{n} \left(y_i - \left(\beta_0 + \beta_1 \times \text{PerCapIncome}_i + \beta_2 \times \text{PopFarms}_i + \beta_3 \times \text{InfantMort}_i\right)\right)^2
\]

```{r}

```

## Problem 2

## True or False Questions

### Question a

**Statement:** The equation \(E[y_i | x_i] = \beta_0 + \beta_1 x_{i1} + \cdots + \beta_p x_{ip}\) is no longer a linear model.

**Answer:** **False**

**Explanation/Justification:**
The equation represents the regression model of \(y_i\) given \(x_i\) as a linear function of the predictor variables \(x_{i1}, x_{i2}, \ldots, x_{ip}\). Despite the presence of \(E\) (the expectation operator), this relationship is still linear in the parameters \(\beta_0, \beta_1, \ldots, \beta_p\). A model is considered linear if it can be expressed as a linear combination of parameters, regardless of the distribution of \(y_i\) or whether we take expectations. Thus, the equation describes a linear regression model.

---

### Question b

**Statement:** If 
\[
A = 
\begin{pmatrix}
4 & 22 & 3 \\
5 & 11 & 2 \\
6 & 0 & 1 
\end{pmatrix}
\]
then the transpose of \(A\) is 
\[
A' = 
\begin{pmatrix}
4 & 5 & 6 \\
22 & 11 & 0 \\
3 & 2 & 1 
\end{pmatrix}
\]

**Answer:** **True**

**Explanation/Justification:**
 The first row of \(A\) becomes the first column of \(A'\), the second row becomes the second column, and the third row becomes the third column. Therefore, the transpose of \(A\) is correctly given by:

\[
A' = 
\begin{pmatrix}
4 & 5 & 6 \\
22 & 11 & 0 \\
3 & 2 & 1 
\end{pmatrix}
\]

---

### Question c

**Statement:** If 
\[
a' = 
\begin{pmatrix}
1.5 & 2 & 3 
\end{pmatrix}
\]
and 
\[
b' = 
\begin{pmatrix}
0 & 10 & 11 
\end{pmatrix}
\]
then the inner product of \(a\) and \(b\) is \(\langle a, b \rangle = 53\).

**Answer:** **True**

**Explanation/Justification:**
The inner product (dot product) of two vectors \(a\) and \(b\) is calculated as follows:

\[
\langle a, b \rangle = a_1 b_1 + a_2 b_2 + a_3 b_3
\]

Substituting the values from the vectors:

\[
\langle a, b \rangle = (1.5 \cdot 0) + (2 \cdot 10) + (3 \cdot 11) = 0 + 20 + 33 = 53
\]


---

## Summary

- **a)** False
- **b)** True
- **c)** True 