---
title: "SET_8"
author: "Oluwanifemi"
date: "2024-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(faraway)
library(ggplot2)
library(alr4)
data("UN11")
#head(UN11)
```



```{r}
model_1 <- lm(formula = lifeExpF ~ I(log(ppgdp)) + pctUrban + fertility,
data = UN11)

summary(model_1)
```

### Question 1(b)

### Conclusion for the Overall F-Test

From the R output, the **overall F-test** evaluates whether at least one of the regression coefficients (excluding the intercept) is significantly different from zero.

- **F-statistic**: 190.5  
- **Degrees of Freedom (DF)**: 3, 195  
- **p-value**: < 2.2 × 10<sup>-16</sup>  
- **Significance Level ($\alpha$)**: 0.01  

#### Conclusion:
The p-value for the F-test (< 2.2 × 10<sup>-16</sup>) is much smaller than the significance level of 0.01.  
**We reject the null hypothesis**, concluding that the model as a whole is statistically significant.

#### Reasoning:
At least one predictor in the model significantly contributes to explaining the variability in life expectancy.

```{r}

reduced_model <- lm(lifeExpF ~ fertility, data = UN11)

# Perform the nested F-test
anova_result <- anova(reduced_model, model_1)
anova_result


```

### Testing Whether the Parameters for `log(ppgdp)` and `pctUrban` Are Both Zero

#### 1. Hypotheses
- **Null Hypothesis (\(H_0\))**: The parameters for `log(ppgdp)` and `pctUrban` are both 0. These variables do not contribute to explaining the variability in `lifeExpF`.
- **Alternative Hypothesis (\(H_a\))**: At least one of the parameters for `log(ppgdp)` or `pctUrban` is not 0. Including these variables significantly improves the model.

#### 2. Test Statistic
The test statistic is the F-statistic from the ANOVA output:
\[
F = 25.814
\]
- **Degrees of Freedom**:
  - Numerator (\(df_1\)): \(2\) (number of predictors added: `log(ppgdp)` and `pctUrban`).
  - Denominator (\(df_2\)): \(195\) (residual degrees of freedom from the full model).

#### 3. Significance Level
The significance level (\(\alpha\)) is \(0.05\).

#### 4. p-value
From the ANOVA output:
\[
p = 1.133 \times 10^{-10}
\]

#### 5. Decision
Compare the p-value to the significance level (\(\alpha = 0.05\)):
\[
p = 1.133 \times 10^{-10} < 0.05
\]
Since the p-value is much smaller than 0.05, we **reject the null hypothesis**.

#### 6. Conclusion
At the 0.05 significance level, there is sufficient evidence to conclude that the parameters for `log(ppgdp)` and/or `pctUrban` are not 0. Adding these predictors significantly improves the model's ability to explain `lifeExpF`.
